---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Travis-CI Build Status](https://travis-ci.org/yutannihilation/gghighlight.svg?branch=master)](https://travis-ci.org/yutannihilation/gghighlight)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "images/"
)
```

# gghighlight

Highlight points and lines in ggplot2.

## Installation

You can install gghighlight from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("yutannihilation/gghighlight")
```

## Example

Suppose the data has a lot of series.

```{r example-data, message=FALSE}
library(dplyr, warn.conflicts = FALSE)

set.seed(1)
d <- tibble(
  idx = 1:10000,
  value = runif(idx, -1, 1),
  type = sample(letters, size = length(idx), replace = TRUE)
) %>%
  group_by(type) %>%
  mutate(value = cumsum(value)) %>%
  ungroup()
```

It is difficult to distinguish them by colour.

```{r ggplot-too-many}
library(ggplot2)

ggplot(d) +
  geom_line(aes(idx, value, colour = type))
```

So we are motivated to highlight only important series, like this:

```{r gghighlight-line}
library(gghighlight)

gghighlight_line(d, aes(idx, value, colour = type), max(value) > 20)
```

As `gghighlight_*()` returns a ggplot object, it is customizable just as we usually do with ggplot2. (Note that, while gghighlights doesn't require ggplot2 loaded, ggplot2 need to be loaded to customize the plot)

```{r gghighlight-line-theme}
gghighlight_line(d, aes(idx, value, colour = type), max(value) > 20) +
  theme_minimal()
```

The plot can be facetted:

```{r gghighlight-line-facet}
gghighlight_line(d, aes(idx, value, colour = type), max(value) > 20) +
  facet_wrap(~ type)
```

### Supported geoms

#### Line

(As shown above)

#### Point


```{r gghighlight-point}
set.seed(10)
d2 <- sample_n(d, 20)

gghighlight_point(d2, aes(idx, value), value > 0)
```
