---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# gghighlight

[![Travis-CI Build Status](https://travis-ci.org/yutannihilation/gghighlight.svg?branch=master)](https://travis-ci.org/yutannihilation/gghighlight)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/gghighlight)](https://cran.r-project.org/package=gghighlight)

Highlight geoms in ggplot2.

## Installation

```{r gh-installation, eval = FALSE}
install.packages("gghighlight")

# Or the development version from GitHub:
# install.packages("devtools")
devtools::install_github("yutannihilation/gghighlight")
```

## Example

Suppose we have a data that has so many series that it is hard to identify them by their colours as the differences are so subtle.

```{r data, include=FALSE}
set.seed(2)
d <- purrr::map_dfr(
  letters,
  ~ data.frame(
      idx = 1:400,
      value = cumsum(runif(400, -1, 1)),
      type = .,
      flag = sample(c(TRUE, FALSE), size = 400, replace = TRUE),
      stringsAsFactors = FALSE
    )
)
```

```{r ggplot2-simple}
library(ggplot2)

ggplot(d) +
  geom_line(aes(idx, value, colour = type))
```

With `gghighlight()`, we can highlight the lines whose max values are larger than 20:

```{r gghighlight-simple}
library(gghighlight)

p <- ggplot(d) +
  geom_line(aes(idx, value, colour = type)) +
  gghighlight(max(value) > 19)

p
```

The result is an usual ggplot object, so it is fully customizable. For example, it can be used with custom themes and facets.

```{r gghighlight-theme-facets}
p + theme_minimal()

p + theme_minimal() + facet_wrap(~ type)
```



`gghighlight()` can highlight almost any geoms. For more details, please read [Introduction to gghighlight](https://yutannihilation.github.io/gghighlight/docs/articles/gghighlight.html).

