---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Travis-CI Build Status](https://travis-ci.org/yutannihilation/gghighlight.svg?branch=master)](https://travis-ci.org/yutannihilation/gghighlight)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/gghighlight)](https://cran.r-project.org/package=gghighlight)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# gghighlight

Highlight lines and points in ggplot2.

## Installation

```{r gh-installation, eval = FALSE}
install.packages("gghighlight")

# Or the development version from GitHub:
# install.packages("devtools")
devtools::install_github("yutannihilation/gghighlight")
```

## Example

Suppose the data has a lot of series.

```{r example-data, message=FALSE}
library(dplyr, warn.conflicts = FALSE)

set.seed(1)
d <- tibble(
  idx = 1:10000,
  value = runif(idx, -1, 1),
  type = sample(letters, size = length(idx), replace = TRUE)
) %>%
  group_by(type) %>%
  mutate(value = cumsum(value)) %>%
  ungroup()
```

It is difficult to distinguish them by colour.

```{r ggplot-too-many}
library(ggplot2)

ggplot(d) +
  geom_line(aes(idx, value, colour = type))
```

So we are motivated to highlight only important series, like this:

```{r gghighlight-line}
library(gghighlight)

gghighlight_line(d, aes(idx, value, colour = type), max(value) > 20)
```

As `gghighlight_*()` returns a ggplot object, it is customizable just as we usually do with ggplot2. (Note that, while gghighlights doesn't require ggplot2 loaded, ggplot2 need to be loaded to customize the plot)

```{r gghighlight-line-theme}
gghighlight_line(d, aes(idx, value, colour = type), max(value) > 20) +
  theme_minimal()
```

The plot also can be facetted:

```{r gghighlight-line-facet}
gghighlight_line(d, aes(idx, value, colour = type), max(value) > 20) +
  facet_wrap(~ type)
```

### Supported geoms

#### Line

```{r ref.label='gghighlight-line'}
```

#### Point

```{r gghighlight-point}
set.seed(10)
d2 <- sample_n(d, 20)

gghighlight_point(d2, aes(idx, value), value > 0)
```

### Grouped vs ungrouped

You may notice that the `gghighlight_line()` and `gghighlight_point()` has different semantics.

By default, `gghighlight_line()` calculates `predicate` per group, more precisely, `dplyr::group_by()` + `dplyr::summarise()`. So if the predicate expression returns more than one value per group, it ends up with an error like this:

```{r grouped-error, error=TRUE}
gghighlight_line(d, aes(idx, value, colour = type), value > 20)
```

On the other hand, `gghighlight_point()` calculates `predicate` per row by default. This behaviour can be controled via `use_group_by` argument like this:

```{r grouped_point}
gghighlight_point(d2, aes(idx, value, colour = type), max(value) > 0, use_group_by = TRUE)
```

While `gghighlight_line()` also has `use_group_by` argument, I don't think ungrouped lines can be interesting because data that can be represented as line must have its series, or groups.

#### Non-logical predicate

To construct a predicate expression like bellow, we need to determine a threshold (in this example, `20`). But it is difficult to choose a nice one before we draw plots.

```{r predicate-example, eval=FALSE}
max(value) > 20
```

So, `gghighlight_*()` allows predicates that return numeric (or character) results. The values are used for sorting data and the top `max_highlight` of rows/groups are highlighted:

```{r numeric-highlight}
gghighlight_line(d, aes(idx, value, colour = type), max(value), max_highlight = 5L)
```
